---
title: "DATA 613 Week 4 Lab 7"
author: "Brenda Cuellar"
date: "2022-07-20"
output: html_document
---
1) Use and show data.table code to select the variables year, month, day, and hour from the imported flights data
```{r}
library(tidyverse)    
library(dplyr)  
library(data.table)
## set wd first
setwd("~/Desktop/R_files/DATA_613")

# Using a data.table read in.
flights <- fread("nycdata.csv")  
flights   
```
```{r}
flights[, list(year, month, day, hour)]->flights1
flights1
```

2) Use and show data. table code to produce a table that shows a carrier of DL, an origin of JFK and a destination of SEA
```{r}
flights[carrier == "DL"& origin == "JFK"& dest == "SEA"]->flights2
flights2
```

3) Use and show data.table code to produce a table that shows a carrier of UA, a month of March, and an airtime that is below 330.
```{r}
flights[carrier == "UA" & month == 3 & air_time < 330]->flights3
flights3
```

4) Use and show tidyverse code to produce a table that shows a carrier of UA, a month of March, and an airtime that is below 330.  
```{r}
flights%>%
  select(carrier, month, air_time)%>%
  filter(carrier == "UA", month == 3, air_time < 330)->flights4
flights4
```

5) Use the data.table method to add a variable called speed that is the average air speed of the plane in miles per hour. 
```{r}
flights[, "speed" := .(distance/air_time)]
flights
```

6) Use the tidyverse method to add a variable called speed that is the average air speed of the plane in miles per hour.
```{r}
flights%>%
  mutate(speed=(distance/air_time)) -> flights6
flights6
```

7) Show and use coding to change the carrier abbreviation of UA to UniitedAir,
7a) data.table method
```{r}
flights[carrier == "UA", carrier := "UnitedAir"][carrier == "UnitedAir"]
```

7b tidyverse method  (Use a sequence of dplyr commands so that you can see the change in your table)
```{r}
flights %>%
  mutate(carrier = recode(carrier, "UA" = "UnitedAir"))%>%
  filter(carrier == "UnitedAir")-> flights7b
flights7b
```